Geometric Signs on Paleolithic Mobile Objects as Cultural Group Markers

### 2025-01-15 endofquarterpaper.qmd

We still lack a complete knowledge of the cultural diversity and movement patterns of the early modern humans from this era, despite a great deal of research into the subject. In this paper, we analyze the distribution and variation of geometric sign types found on mobile artefacts from the Upper Paleolithic in Europe during the Aurignacian technocomplex. The basis of this analysis is the idea that variations on signs can serve as an indicator of group membership and cultural similarity. This approach is supported by both ethnographic evidence throughout history, as well as similar archaeological studies done on ornamental and parietal art [@Pereira_Manen_2023; @Sauvet_2018; @Rigaud_dErrico_Vanhaeren_2015; @baker2024evidence; @Lycett_2019; @Vanhaeren_d]. In this study, we use our analyses to attempt to answer the following questions: Do identifiable and relevant patterns of cultural groups exist in the data, and if so, what can they tell us about the patterns of diversity and movement among Upper Paleolithic hominids?

<!--SC: I'm still having a difficult time figuring out just how to end this paragraph - you mentioned that I should have a research question ending in a question mark, but I'm struggling to figure out how to phrase it in a way that is appropriately formal  -->

>>>>

We still lack a complete knowledge of the cultural diversity and movement patterns of the early modern humans from this era, despite a great deal of research into the subject. In this paper, we analyze the distribution and variation of geometric sign types found on mobile artefacts from the Upper Paleolithic in Europe during the Aurignacian technocomplex. The basis of this analysis is the idea that variations on signs can serve as an indicator of group membership and cultural similarity. This approach is supported by both ethnographic evidence throughout history, as well as similar archaeological studies done on ornamental and parietal art [@Pereira_Manen_2023; @Sauvet_2018; @Rigaud_dErrico_Vanhaeren_2015; @baker2024evidence; @Lycett_2019; @Vanhaeren_d]. In this study, we address the following questions: Are cultural groups evident in the distribution of geometric signs, and if so, what can they tell us about the cultural diversity and movement of Upper Paleolithic humans?

<!--BM: our position is that cultural groups exist, our contribution is to see if/how they are evident in the sign data  -->
